--- 
title: "Rock Your Resume with R"
author: "Mai Uchida"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
---

# Welcome! {-}

<iframe width='100%' height='800' src='https://rdrr.io/snippets/embed/' frameborder='0'></iframe>


```{r eval=FALSE, include=FALSE}
install.packages('bookdown')
options(max.print = 100)
knitr::opts_chunk$set(message=FALSE, fig.path="R/Rintro/figures/")
knitr::opts_knit$set(root.dir="R/Rintro") # base.url = "/" # base.dir="R/Rintro"
# knitr::opts_chunk$set(eval=FALSE, results=FALSE, message=FALSE, warning=FALSE, error=FALSE)
# knitr::opts_knit$set(base.dir = "/Users/nunnlab/Desktop/GitHub/rgriff23.github.io/")
```


```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

# Why learn R?

## Pros and Cons of R {-}

Pros

  * R is a language made by statisticians
     + Lots of packages that allow you to perform statistical analysis/ modeling  
     + Large user community  
     + Free and open source  
     + Nice visualization package (ggplot) 
    
Cons 

  * R has a learning curve  
  * R is more specific to statistical analysis 
     + (e.g.) Python can do a lot more and their stats packages are getting better
     
## Learning a programing language is an investment, choose wisely {-}

Points of considerations 

  * What’s your industry’s standard?  
      + If you are going to finance, I’d learn Excel instead
      + If a job you are applying to uses SAS, I'd learn SAS instead
  <br />
  <br />
  * Do you need a programming language to perform what you need to do? 
      + If you want to a health insurance company writing policies, maybe learning a programming language will have a small investment of return 
  <br />
  <br />
  * Is the software/ programming language free (open source) or licensed?  
      + If licensed and your employer/ research team has no funding for it, your programing knowledge won’t be as helpful as it could be  
      + If open source, you can just download it to your computer (most of the time if no restrictions from the organizations you work for) 
      
# Why visualization matters?

# Difference between exploratory and explanatory data visualization

# Base R syntax and data type

## Data type
  + Charactor vector (srting and/ or factor): defined by using " "
    + (e.g.) "Jack", "Rose"
    + Cannot perform mathmatical calculations
    
  + Numeric vector (double, integer etc...): numbers are written as is
    + (e.g.) 100, 200 
    + Can perfom a mathmatical calculations
  <br />
  + Logical vecor: TRUE or FALSE
      
Let's practice!

  * write a code *100 + 200* as a numerical vector
  * write a code *100 + 200* as a character vector

<iframe width='100%' height='400' src='https://rdrr.io/snippets/embed/?code=' frameborder='0'></iframe>
  

## Create a variable and assign a value 
* if you want to make a variable x and assign 1 as a value
    + "<-" is the symbol we use to assign values in R
    + "#" is to make comments on an R script
```{r, eval=FALSE, include=TRUE, message=FALSE}
x <- 1
x #check the variable
```

Let's practice!

  * Create a variable pineapples and assign 120 as a value
      + Check the variable
  
<iframe width='100%' height='400' src='https://rdrr.io/snippets/embed/?code=' frameborder='0'></iframe>

## Create variables and calculate
* Create variables x and y. 
* Assign 1 as a value for x, 4 as a value of y
* Add them up
* Check if x is bigger than y

```{r, eval=FALSE, include=TRUE, message=FALSE}

#create variables x, y and assign values
x <- 1
y <- 4

#add x and y
x + y

#examine using logical operator 
x > y

```
Let's practice!

* Create a variable pineapples and assign 120 as a value. 
* Create a variable strawberries and assign 70 as a value.
* Add them up using variable names
* Check if pineapples are larger than strawberries using a logical operator

<iframe width='100%' height='400' src='https://rdrr.io/snippets/embed/?code=' frameborder='0'></iframe>


## Create a new variable with the sum 
* Create a variable z and assign the sum of x + y
```{r, eval=FALSE, include=TRUE, message=FALSE}
x <- 1
y <- 4

#create a new variable z and assign the sum of x and y
z <- x + y

#check the value of z
z
``` 
Let's practice!

* Create a variable pineapples and assign 120 as a value. 
* Create a variable strawberries and assign 70 as a value.
* Create a new variable pineberries and assign the sum of pineapples and strawberries
* Check the value of pineberries)

<iframe width='100%' height='400' src='https://rdrr.io/snippets/embed/?code=' frameborder='0'></iframe>

## Use functions
* Calcurate the sum of x, y and z 
```{r, eval=FALSE, include=TRUE, message=FALSE}
x <- 1
y <- 4
z <- x + y

#sum up x, y and z using the sum() function
sum(x, y, z)
```
Let's practice!

* Create a variable pineapples and assign 120 as a value. 
* Create a variable strawberries and assign 70 as a value.
* Create a new variable pineberries and assign the sum of pineapples and strawberries
* Sum up pinabbles, strawberries, and pineberries using the sum() function

<iframe width='100%' height='400' src='https://rdrr.io/snippets/embed/?code=' frameborder='0'></iframe>

## Make a data frame
* Make a data frame "df1"
    + (explanation 1). "variable" is a column name, c("", "",) is to add multiple vairiables. this says assign x to the "variable" column,  row 1. assign y to the "variable" column, row 2. assign z to the "variable" column, row 3. 
    + (explanation 2). "value" is a column name. this says that assign 1 to the "value" column, row 1. assign 4 to the "value" column, row 2. assign 10 to the "value" column, row 3. 
* Calcurate the mean of x, y and z 
```{r, eval=FALSE, include=TRUE, message=FALSE}
df1 <- data.frame("variable" = c("x", "y", "z"), #(explanation 1). 
                   "value" = c(1, 4, 10)) #(explanation 2). 

str(df1)

#calcurate mean
mean(df1$value) #"$" is select "value column" from df1
```

Let's practice!

* Create a dataframe "fruits," which contains two columns ("fruitType" and "amount")
    + pineapples, which the value is 120
    + strawberries, which the value is 70
    + pineberries, which the value is 190
* Check the structure of the dataframe 
* Calculate the mean of "amount"" 
    
<iframe width='100%' height='400' src='https://rdrr.io/snippets/embed/?code=' frameborder='0'></iframe>

## Create a simple plot

* Create a data frame "firstHalf"
* Create a simple barplot using the "barplot()" function
```{r, eval=FALSE, include=TRUE, message=FALSE}
firstHalf <- data.frame("month" = c("Jan", "Feb", "Mar", "Apr", "May", "June"), 
                   "value" = c(150, 200, 15, 70, 300, 60)) 

#create a bar plot. names.arg is to label each bar
barplot(fristHalf$value, names.arg = c("Jan", "Feb", "Mar", "Apr", "May", "June")) 
```

Let's practice!

* Create a data frame "secondHalf." Monthly values are as follows: 
    + July, 250
    + August, 10
    + September, 600
    + October, 120
    + November, 80
    + December, 50

* Create a barplot using the "barplot()" function
    + Make sure to label each bars
    
<iframe width='100%' height='400' src='https://rdrr.io/snippets/embed/?code=' frameborder='0'></iframe>

# Practice Problem 1

## Let’s make Anscombe's quartet {-}

What is Anscombe's Quartet? 



General Steps for Data Analysis

1. Load data (of course, you need to have your data prior to this...)
2. Clean data 
3. Check data distributions
4. Data frame creation (subsetting the data)
5. Check data distribution 
    + Summary statistics 
    + Exploratory data visualization
6. Analyze your data 
7. Create a explanatory data visualization product


## Set up an environement and load data
  * Install R and R studio 
      + There are a couple options for R's Integrated Development Environment (IDE), but I would recommend R Studio if you do not have your preferred platform yet. 
  <br />
  <br />
  * Create an R peoject file (optional, but recommended)
      + name your file: myFirstR
  <br />
  <br />
  * Create an R script file
      + name your file: anscombe
  <br />
  <br />
  * Check and set up your work directory 
      + work directory: a place/ folder on your computer where your file locates
```{r, include=TRUE, eval=FALSE, message = FALSE}
#check your work directory
getwd()

#set your work directory 
setwd("/YOUR FILE PATH HERE/")

#mac users (if you made a folder on your desktop)
setwd("/Users/your-user-name/Desktop/myFirstR")

#PC users (if you made a folder on your desktop)
setwd("\\")
```

  * Install and load packages you need for the project
      + What are packages?
      + Difference between install and load packages 
          + Install packages: You only need to do it once when you don't have a package installed in your R
          + Load packages: You need to load them when you re-open R or want to use a package that is not loaded yet (package is already installed)
          + Make sure you install and load all the packages you need in the beginning of your code 
```{r, include=TRUE, eval=FALSE, message=FALSE}
#install packages 
install.packages("ggplot2")
install.packages("dyplyr")
install.packages("tidyverse")
install.packages("RCurl")
  
#load packages
library(ggplot2)
library(dplyr)
library(tidyverse)
library(RCurl)
```

```{r, include=TRUE, eval=FALSE, message=FALSE}
#to check installed packages, run the code below or click "packages" on R studio
ip = as.data.frame(installed.packages()[,c(1,3:4)])
ip = ip[is.na(ip$Priority),1:2,drop=FALSE]ip 
```
        
* load a dataset (if you are running R on R studio write this code to load data)
```{r, include=TRUE, eval=FALSE, message=FALSE}
#loading data from a github repository. assign URL variable
url <- getURL("https://raw.githubusercontent.com/maiuchi/r_test1/master/test.csv")

#name a data frame (df) anscombe and read csv
anscombe <- read.csv(text = url)
```

* R already has this data installed... so you can also do... 
```{r, include=TRUE, eval=FALSE, message=FALSE}
data("anscombe")
```

## Check data structure and clean data (done for today)

* Once you load the data up to R, the first thing you'd want to do is: 
    + Make sure your data is properly loaded 
        + function to use: View(name-of-dataframe)
    + Check the structure of the data
        + function to use: str(name-of-dataframe) or glimpse(name-of-dataframe)

## Check data distributions

# Temporaly R Platform {-}

In case you had issues running code on R studio, you can use type your code below to follow the demo

<iframe width='100%' height='1800' src='https://rdrr.io/snippets/embed/?code=' frameborder='0'></iframe>
